---
title: "Chaining concordance tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chaining concordance tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(knitr.kable.NA = ".")
```

```{r include=FALSE}
library(kableExtra)
library(dplyr)
```

`comtradeRggregator` provides an option for easy reclassification of trade data using official concordance tables from the [United Nations Statistical Division (UNSD)](https://unstats.un.org/unsd/trade/classifications/correspondence-tables.asp) and the [World Integrated Trade Solution (WITS)](https://wits.worldbank.org/product_concordance.html); see [Concordance Table](https://amannj.github.io/resources/comtradeRggregator/index.html#concordance-table) below. To access the respective concordance tables in R, type `<tradecode from>_<tradecode to>`. For example, the concordance table for reclassifying `H3` trade data according to the `H0` nomenclature is stored as object `H3_H0` and its R help file can be accessed by typing `?H3_H0`.

However, in some cases no official concordance tables between two standards are available. For example, there is no direct concordance for reclassifying `H4` trade data to `MTN`.

```{r concordanceTable, echo=FALSE}

hs <- c("HS", "W", NA, "W", "W", "W", "W", "W", "W", "W", "W", "W", "W", NA, NA, "W", "W", "W", "W", "W", "W")
h0 <- c("H0", "W", NA, "W", "W", "W", NA, NA, NA, NA, NA, "W", "W", NA, NA, "W", "W", "W", "W", "W", "W")
h1 <- c("H1", "W", NA, "W", "W", "W", "W/U", NA, NA, NA, NA, "W", "W", NA, NA, "W", "W", "W/U", "W/U", "W/U", "W")
h2 <- c("H2", "W", NA, "W", "W", "W", "W/U", "W/U", NA, NA, NA, "W", "W", NA, NA, "W", "W", "W/U", "W/U", "W/U", "W")
h3 <- c("H3", "W", NA, "W", "W", "W", "W/U", "W/U", "W/U", NA, NA, "W", "W", NA, NA, "W", "W", "W/U", "W/U", "W/U", "W/U")
h4 <- c("H4", "U", "U", NA, "U", "U", "U", "U", "U", "U", NA, NA, NA, "U", "U", NA, NA, "U", "U", "U", "U")
h5 <- c("H5", "U", NA, NA, NA, NA, "U", "U", "U", "U", "U", NA, paste0("x", footnote_marker_alphabet(1)), NA, NA, NA, NA, "U", "U", "U", "U")

s1 <- c("S1", "W", rep(NA, 19))
s2 <- c("S2", "W/U", rep(NA, 9), "W", NA, NA, NA, NA, NA, "U", NA, NA, NA)
s3 <- c("S3", rep(NA, 11), NA, NA, NA, NA, NA, "U", "U", NA, NA)
s4 <- c("S4", rep(NA, 11), NA, NA, NA, NA, NA, NA, NA, NA, NA)

i2 <- c("I2", rep(NA, 11), "U", "U", rep(NA, 7))
i3 <- c("I3", rep(NA, 10), "U", NA, "U", rep(NA, 7))
i31 <- c("I3.1", rep(NA, 11), "U", NA, "U", rep(NA, 6))

df <- rbind(hs, h0, h1, h2, h3, h4, h5, s1, s2, s3, s4, i2, i3, i31)

table <- tibble(
  df
)

kable(table,
  caption = "Table 1: Concordance Table",
  col.names = c(
    "",
    "BEC4", "BEC5", "CCCN", "CPC", "GTAP",
    "H0", "H1", "H2", "H3", "H4",
    "I2", "I3", "I31", "I4",
    "MTN", "SIC",
    "S1", "S2", "S3", "S4"
  ),
  escape = F
) %>%
  kable_styling(font_size = 8) %>%
  add_header_above(c(" " = 1, "BEC" = 2, " " = 3, "Harmonised System" = 5, "ISIC" = 4, " " = 2, "SITC" = 4)) %>%
  add_header_above(c("From...", "...to" = 20)) %>%
  pack_rows("Harmonised System (HS)", 1, 7) %>%
  pack_rows("Standard International Trade Classification (SITC)", 8, 11) %>%
  pack_rows("International Standard Industrial Classification (ISIC)", 12, 14) %>%
  footnote(
    general = " W for WITS; U for UN-Stats.",
    general_title = "Note: ",
    alphabet = c(
      ": Example for concordance across seemingly unrelated classifications: H5 >> H3 >> I3."
    ),
    alphabet_title = "x",
    # symbol = "Reclassification of SITC trade data requires data on aggregation levels four (AG4) and five (AG5) together.",
    # symbol_title = "Note: ",
    # paste0("S1", footnote_marker_symbol(1))
    footnote_as_chunk = T
  )
```

-------------

# Practical examples

It goes without saying that in an ideal world, one would prefer to download and convert trade data for which there exists an official concordance table, as this will always give you identical or better results than when chaining concordance tables. There are, however, situations, where this might not be possible, and we need to resort to the chaining of concordance tables to get from a  
particular desired classification to another one.

The subsequent section should therefore be read as a set of (not comprehensive) examples of how to chain concordance tables using `comtradeRggregator` without any guarantee that this will (always) produce sensible results.

```{r message=FALSE}
library(comtradeRggregator)
```

----------------

# Concordance `H4` to `MTN`

For this example, we start with a data extract following the `HS 2012` (`H4`) nomenclature, which we want to convert to the `Multilateral Trade Negotiations` (`MTN`) standard. As illustrated in the [Concordance Table](https://amannj.github.io/resources/comtradeRggregator/index.html#concordance-table) above, there is no designated concordance table available. We can, however, use the fact that

- there is a designated concordance table between `H4` and `H3`;
- there is another designated concordance table between `H3` and `MTN`.

Consequently, we can create our own `H4` to `MTN` classification by following the steps below:

1. Download Comtrade data using `download_Comtrade()`;
2. Use `convert_Comtrade()` to convert trade data from `H4` to `H3`;
3. Use `convert_Comtrade()` again to convert from `H3` to `MTN`.

### Step by step

#### Step 1. Download Comtrade data using `download_Comtrade()`

Let's again assume we want to download `annual` global, i.e. `World`,  `commodity`  trade data in `USD` following the `HS2012` nomenclature for `exports` at `AG6` level  reported by `Austria`  in `2018`:

```{r message=FALSE, warning=FALSE}
AT_World_H4 <- download_Comtrade(
  year = "2018",
  frequency = "annual",
  countries = "Austria",
  partners = "World",
  nomenclature = "HS2012",
  commodity = "AG6",
  type = "commodities",
  select.stats = "trade_value_usd",
  direction = "exports"
)
```

#### Step 2. Use `convert_Comtrade()` to convert trade data from `H4` to `H3`

```{r}
AT_World_H4_H3 <- AT_World_H4 %>%
  convert_Comtrade(
    classification.from = "H4",
    classification.to = "H3"
  )
```

As you can see, `convert_Comtrade()` informs you that commodity code `999999 - Commodities not specified according to kind` contained in `AT_World_H4` could not be matched to any `H3` classification. In the given example, this is fine as this particular commodity code belongs to a "catch-all" category of otherwise unspecified commodities:

```{r}
AT_World_H4_H3 %>%
  filter(commodity_code == "999999") %>%
  select(commodity_code, commodity, `HS 2007 Product Code`, trade_value_usd)
```

We lose this "catch-all" category during the conversion step, and you should always double-check to make sure you do not 'lose too many' commodity groups when converting between standards.

#### Step 3. Use `convert_Comtrade()` again to convert from `H3` to `MTN`

```{r}
AT_World_H4_H3_MTN <- AT_World_H4_H3 %>%
  convert_Comtrade(
    classification.from = "H3",
    classification.to = "MTN"
  )
```

Here you can see that in the last step, `convert_Comtrade()` has more difficulties matching the respective commodity codes to `MTN` aggregates. 

Is this problem the result of chaining concordance tables? We explore this in the next section.

### How good is the `H4` to `MTN` concordance?

In the table below, we compare the results of the different conversion steps where we look at the trade volume of exports (in USD) of  `AT_World_H4` (the baseline) vs. `AT_World_H4_H3` and `AT_World_H4_H3_MTN`. We want to know how much of the baseline data is lost during the different conversion steps. The two metrics for this are 

- the absolute difference between the trade volume of exports (in USD) after each reclassification vs. the baseline; and
- the percentage difference of the same aggregates.

As we want to compare the quality of the results when using chained concordance tables (and not the quality of the official concordance table between `H3` and `MTN`), we also download `H3` trade data directly and convert it to `MTN` using official concordance tables only (i.e. no chaining of concordance tables).

```{r message=FALSE, warning=FALSE}
AT_World_H3_MTN <- download_Comtrade(
  year = "2018",
  frequency = "annual",
  countries = "Austria",
  partners = "World",
  nomenclature = "HS2007",
  commodity = "AG6",
  type = "commodities",
  select.stats = "trade_value_usd",
  direction = "exports"
) %>%
  convert_Comtrade(
    classification.from = "H3",
    classification.to = "MTN"
  )
```

As Table 2 below illustrates, the conversion from `H4` to `H3` results in a loss of roughly five percentage points of trade data. The difference in the sums is exclusively due to `999999 - Commodities not specified according to kind` as established before. Chaining concordance tables results in a total loss of 8 percentage points which is indeed notable. However, as `AT_World_H3_MTN` indicates, this loss is mostly due to the unmapped category `999999 - Commodities not specified according to kind`, and not the result of chaining concordance tables.

In other words, chaining concordance tables to transform `H4` trade data to the `MTN` classification does seem to produce sensible results based on the explored example.

```{r echo=FALSE, message=FALSE, warning=FALSE}
AT_World_H4 %>%
  summarise(AT_World_H4 = sum(trade_value_usd)) %>%
  pull() -> c11

AT_World_H4_H3 %>%
  filter(!is.na(`HS 2007 Product Code`)) %>%
  summarise(AT_World_H3 = sum(trade_value_usd)) %>%
  pull() -> c12

AT_World_H4_H3_MTN %>%
  filter(!is.na(`MTN Categories Product Code`)) %>%
  summarise(AT_World_H4 = sum(trade_value_usd)) %>%
  pull() -> c13

AT_World_H3_MTN %>%
  filter(!is.na(`MTN Categories Product Code`)) %>%
  summarise(AT_World_H3_MTN = sum(trade_value_usd)) %>%
  pull() -> c14


tibble(
  Summary = c("Trade value (in USD)", "Absolute difference to `AT_World_H4`  (in USD)", "Share in `AT_World_H4` (%)"),
  AT_World_H4 = c(c11, abs(c11 - c11), round(c11 / c11 * 100, 0)),
  AT_World_H4_H3 = c(c12, abs(c12 - c11), round(c12 / c11 * 100, 0)),
  AT_World_H4_H3_MTN = c(c13, abs(c13 - c11), round(c13 / c11 * 100, 0)),
  AT_World_H3_MTN = c(c14, abs(c14 - c11), round(c14 / c11 * 100, 0))
) -> df_out


df_out %>%
  mutate_if(is.numeric, scientific = F, format, digits=0) %>%
  kable(
    caption = "Table 2: Comparison concordance results",
    col.names = c("", "`AT_World_H4`", "`AT_World_H4_H3`", "`AT_World_H4_H3_MTN`", "`AT_World_H3_MTN`")
  ) %>%
  kable_styling()
```
