---
title: "Extracting aggregated trade data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting aggregated trade data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Trade data is often reclassified following different standards; see [Concordance Table](https://amannj.github.io/resources/comtradeRggregator/index.html#concordance-table)
for all mappings that are possible using `comtradeRggregator`.

However, in all those cases the internal logic of the conversion is that we use
'all' trade data, which is then mapped onto a different classification. What if, 
in turn, we want to extract only a set of traded commodities that belong to only
a specific sub-group of the target classification? 

In the example below we assume we want to extract all *SITC Rev. 2* (`S2`) trade data
that belongs to a very specific manufacturing sector, say, *ISIC Rev. 2* (`I2`)
sector `3530 - Petroleum refineries`. We can accomplish this in a few simple
steps:



First, we load the `comtradeRggregator` package.

```{r message=FALSE}
library(comtradeRggregator)
```


Next, we extract all `S2` codes that correspond to the `I2` sector `3530`
which we want to extract:

```{r message=FALSE, warning=FALSE}
comm_petrol <- S2_I2 %>%
  filter(`ISIC Revision 2 Product Code` == 3530)
comm_petrol
```

In a next step we use the SITC commodity codes of `comm_petrol` as an input argument
for the API query:

```{r message=FALSE, warning=FALSE}
data_extract <- download_Comtrade(
  year = "2018",
  frequency = "annual",
  countries = "France",
  partners = "Germany",
  nomenclature = "S2",
  commodity = comm_petrol %>% pull(`SITC Revision 2 Product Code`),
  select.stats = "trade_value_usd"
)
```


Finally, we re-classify the extracted data back to `S2` using 
`convert_Comtrade()`:

```{r}
convert_Comtrade(data_extract,
                 classification.from = 'S2',
                 classification.to = 'I2') %>%
  distinct(`ISIC Revision 2 Product Code`)

```
