---
title: "Working with mirror data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling error messages}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## A more complex example


Let's say you want to download monthly mirror trade data starting in 2019 at 
the HS4 6-digit level for a handful of countries.


While there are multiple areas where both comtrade and WITS offer similar functionality
to the comtradeRggregator package, this is a scenario where both alternatives
are tedious to work with:

- While WITS does offer a straight-forward integration of mirror data, there are
 (at the time of writing this documentation in 2021) no options available to query 
 monthly data;
 
 - While comtrade does over the most recent and granular trade data, the options 
 to query mirror data are limited.


As I will demonstrate below, this data can be obtained via comtradeRggregator with
little to no additional effort.


## Implementation

First, load the package.


```{r message=FALSE}
library(comtradeRggregator)
```

```{r, include = FALSE}
# Load token; do not show
my_token <- Sys.getenv("COMTRADE_TOKEN")
my_location <- "C:/Users/amann/Desktop/tmp"
```

Define list of countries for extraction

```{r}
country_list = c("Afghanistan", "Angola")
```

Let's break up this query into two parts: a download for the years 2019 and 2020,  and a separate query for 2021. In such somewhat more elaborate cases,  it is advisable to keep any temporarily downloaded data by setting 
`rm.temporaryFiles = FALSE`. I also want to save the temporary files in a 
different location specified in variable `my_location`, so I also specify 
`location.temporaryFiles = my_location` in the query below. In an ideal world, you will have access to an authenticated account 
[(see here)](https://comtrade.un.org/data/doc/api/). I do, so I'm also providing my comtrade authentication token as part of the query. Finally, as we
do not want to overwhelm the API, we reduce the number of extracted countries per sub-query to three, i.e. `ext_cnt  = 3`, instead of the default of five, 
and increase the cool-down period between the separate queries from five to ten seconds between rounds by setting `speed = 10`. These last two steps are
optional and only serve illustrative purposes.


Note that for the first query we are not specifying 
the parameter `month` which, by default, will return data for all 12 months.


```{r eval=FALSE, message=FALSE, warning=FALSE}
mirr_part1 <- download_Comtrade(
  year = c("2019", "2020"),
  frequency = "monthly",
  countries = country_list,
  partners = "World",
  nomenclature = "HS2012",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  location.temporaryFiles = my_location,
  build.Comtrade = TRUE,
  token = my_token,
  ext_cnt  = 3,
  sleep = 10
)
```


Finally, we want to find the last month of 2021 for which any comtrade data is 
available to download.

You can either use the `is.available_Comtrade()` function for this, or, alternatively,
the data file `Comtrade_DA`, which provides comtrade's data availability as a tibble.
At the time of writing this tutorial, the most recent monthly trade data is July 2021.

```{r}
Comtrade_DA %>% 
  mutate(yr = substr(ps, 1, 4), mnth = substr(ps, 5,6)) %>% 
  filter(freq == "MONTHLY" & yr == '2021' ) %>%
  summarise(x = max(mnth))
```



```{r eval=FALSE, message=FALSE, warning=FALSE}
mirr_part2 <- download_Comtrade(
  year = c("2021"),
  frequency = "monthly",
  month = c("01":"07"),
  countries = country_list,
  partners = "World",
  nomenclature = "HS2012",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  location.temporaryFiles = my_location,
  build.Comtrade = TRUE,
  token = my_token,
  ext_cnt  = 3,
  sleep = 10
)
```
