---
title: "Working with mirror data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling error messages}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Why use mirror data?

Sometimes trade data between countries do not "add up": For example, in 2020 
country A reported USD $X$ millions exports to country B while B was reporting
USD $Y$ millions, where $X \neq Y$; which data is more reliable?
Alternatively, consider the case where country A does not report any trade data 
at all; we can use trade data from all
trade partners of country A to reconstruct its 'own' trade data?

In such cases, it makes sense to use [mirrored trade data](https://wits.worldbank.org/wits/wits/witshelp/Content/Data_Retrieval/T/Intro/B2.Imports_Exports_and_Mirror.htm). 
The function `download_Comtrade()` offers an easy option to accomplish this. While
[Example 4](https://amannj.github.io/resources/comtradeRggregator/articles/comtradeRggregator.html#example-4)
in the *Get Started guide* provides a simple introduction, a more complex
example is discussed in this article.




## A more complex example



Let's say you want to download monthly mirror trade data starting in 2019 at 
the HS 6-digit level for a handful of countries for which there is no trade data
available on [comtrade](https://comtrade.un.org/).

While there are multiple areas where both [comtrade](https://comtrade.un.org/) and 
[WITS](https://wits.worldbank.org/) offer similar functionality
to the `comtradeRggregator` package, this is a scenario where both alternatives
are tedious to work with:

- While [WITS](https://wits.worldbank.org/) does offer a straight-forward integration of mirror data, there are
 (at the time of writing this documentation in 2021) no options available to query 
 monthly data;
 
 - While [comtrade](https://comtrade.un.org/) does over the most recent and granular trade data, the options 
 to query mirror data are not there.


As I will demonstrate below, the same query can be run with
little to no additional effort using `comtradeRggregator`.


## Implementation

First, load the package.


```{r message=FALSE}
library(comtradeRggregator)
```

```{r, include = FALSE}
# Load token; do not show
my_token <- Sys.getenv("COMTRADE_TOKEN")
my_location <- "C:/Users/amann/Desktop/tmp"
```

Next, we want to define list of countries for extraction. The two examples we 
use here are *Afghanistan* and the *Democratic Republic of the Congo*. 
To double-check the correct spelling of the country names we use 
`is.available_Comtrade()` with option `is.fuzzy = TRUE`. Below I run the check 
for the *Democratic Republic of the Congo* only. However, for large queries it
is advisable to double-check that all country names are correct before setting
up the actual data query. If you don't do that, and are unlucky in the sense that
a particular country name is spelled differently in 
[comtrade](https://comtrade.un.org/), your data query will return an error.

```{r message=FALSE, warning=FALSE}
is.available_Comtrade(
  is.contained = "Congo",
  is.fuzzy = TRUE,
  frequency = "annual",
  nomenclature = "HS",
  year = '2019'
)
```

The final list of countries we want to query is:

```{r}
country_list = c("Afghanistan", "Dem. Rep. of the Congo")

```


When setting up this more elaborate query, there are a few things we should
keep in mind:


- It makes sense to break up this query into two parts: a download for the years 2019 and 2020,
and a separate query for 2021. The reason for this is that at the time of writing 
this  tutorial, trade data for all 12 months was available for 2019 and 2020, 
but not for 2021 as this would make `download_Comtrade()`
throw an error message.


- It is advisable to keep any temporarily downloaded data by setting 
`rm.temporaryFiles = FALSE`. I also want to save the temporary files in a 
different location specified in variable `my_location`, so I also specify 
`location.temporaryFiles = my_location` in the query below. 

- In an ideal world, you will have access to an authenticated account 
[(see here)](https://comtrade.un.org/data/doc/api/). 
I do, so I'm also providing my [comtrade](https://comtrade.un.org/) 
authentication token as part of the query. 

- Note that for the first query we are not specifying 
the parameter `month` which, by default, will return data for all 12 months.
Also, note that monthly [comtrade](https://comtrade.un.org/) data is only available following the 
*HS combined nomenclature*.

- Finally, as we do not want to overwhelm the API, we reduce the number of 
extracted countries per sub-query to three, i.e. `ext_cnt  = 3`, instead of
the default of five, and increase the cool-down period between the separate 
queries from five to ten seconds between rounds by setting `speed = 10`. 
These last two steps are optional and only serve illustrative purposes.


With all of this out of the way, the final monthly mirror data query for 2019 and
2020 looks like this:

```{r eval=FALSE, message=FALSE, warning=FALSE}
mirr_part1 <- download_Comtrade(
  year = c("2019", "2020"),
  frequency = "monthly",
  countries = country_list,
  partners = "World",
  nomenclature = "HS",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  location.temporaryFiles = my_location,
  build.Comtrade = TRUE,
  token = my_token,
  ext_cnt  = 3,
  sleep = 10
)
```


For the 2021 query, we first want to find the last month of 2021 for which 
trade data is available. This makes sense because we do not want to run a query for
a month where no trade data is available as this would make `download_Comtrade()`
throw an error message.

To check the data availability, you can either use the 
`is.available_Comtrade()` function, or
the data file `Comtrade_DA` which I am doing below. As you can see, at the 
time of writing this tutorial, the most recent monthly trade data is July 2021.


```{r}
Comtrade_DA %>% 
  mutate(yr = substr(ps, 1, 4), mnth = substr(ps, 5,6)) %>% 
  filter(freq == "MONTHLY" & yr == '2021' ) %>%
  summarise(x = max(mnth))
```

The query for 2021 is identical to the first one with the exception of parameter
`month = c(1:7)`. 

```{r eval=FALSE, message=FALSE, warning=FALSE}
mirr_part2 <- download_Comtrade(
  year = c("2021"),
  frequency = "monthly",
  month = c(1:7),
  countries = country_list,
  partners = "World",
  nomenclature = "HS",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  location.temporaryFiles = my_location,
  build.Comtrade = TRUE,
  token = my_token,
  ext_cnt  = 3,
  sleep = 10
)
```
