---
title: "Working with mirror data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling error messages}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## What is mirror data and why use it?

Consider the case where country A does not report any trade data at all. What do we do?
In such cases it makes sense to use [mirrored trade data](https://wits.worldbank.org/wits/wits/witshelp/Content/Data_Retrieval/T/Intro/B2.Imports_Exports_and_Mirror.htm). In a nutshell, using the mirrored data approach we can reconstruct trade data for country A, even if the country itself does not report andy trade figures, by looking at all trade flows in which country A participated. For example, you can reconstruct country A's exports to country B by looking at country B's imports from country A. 

Do note, however, that this method is not free of limitations. Most importantly, the reliability
of the mirrored trade data for country A depends on the data availability of trade data
made available by A's trading partners and may potentially be systematically biased. Secondly,
as comtrade trade data undergoes constant revisions and updates, changes in the mirrored
aggregate are more likely than with non-mirrored data.

These words of caution aside, the function `download_Comtrade()` offers an easy option perform mirror data extracts.  Below, we first revisit [Example 4](https://amannj.github.io/resources/comtradeRggregator/articles/comtradeRggregator.html#example-4)
from the *Get Started guide* provides a simple introduction to this topic. Next,
we turn to somewhat more involved user cases.



## A simple example

First, load the package.


```{r message=FALSE}
library(comtradeRggregator)
```


In [Example 4](https://amannj.github.io/resources/comtradeRggregator/articles/comtradeRggregator.html#example-4) we reconstructed Austria's export to Germany by looking at Germany's imports from Austria by setting `is.mirrorData = TRUE`. As you can see, variable `reporter` now is returned as `country name + "mirrored"`, indicating that the  the data was derived from a mirrored data query.


```{r eval=FALSE, message=FALSE, warning=FALSE}
mirrX_AT_DE <- download_Comtrade(
  year = "2018",
  frequency = "annual",
  countries = "Austria",
  partners = "Germany",
  nomenclature = "H3",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "exports",
  type = "commodities",
  is.mirrorData = TRUE
)
mirrX_AT_DE
#> # A tibble: 4,056 x 8
#>    classification period trade_flow reporter         partner commodity_code commodity          trade_value_usd
#>    <chr>           <int> <chr>      <chr>            <chr>   <chr>          <chr>                        <dbl>
#>  1 H3               2018 Export     Austria mirrored Germany 010110         Live horses/asses~          578079
#>  2 H3               2018 Export     Austria mirrored Germany 010190         Live horses/asses~           17717
#>  3 H3               2018 Export     Austria mirrored Germany 010210         Live bovine anima~           75642
#>  4 H3               2018 Export     Austria mirrored Germany 010290         Live bovine anima~         4329765
#>  5 H3               2018 Export     Austria mirrored Germany 010391         Live swine other ~          916695
#>  6 H3               2018 Export     Austria mirrored Germany 010392         Live swine other ~          219879
#>  7 H3               2018 Export     Austria mirrored Germany 010410         Live sheep                  778978
#>  8 H3               2018 Export     Austria mirrored Germany 010420         Live goats                   45964
#>  9 H3               2018 Export     Austria mirrored Germany 010511         Live fowls of spe~         6716350
#> 10 H3               2018 Export     Austria mirrored Germany 010512         Live turkeys, wei~            2194
#> # ... with 4,046 more rows
```

Conceptually, reported imports by Germany from Austria and reported mirrored exports by Austria to Germany are identical by definition. Note that in the query below we can use either `tradecode` names or abbreviations as input arguments for `tradecode` as provided in table [Trade Classifications](https://amannj.github.io/resources/comtradeRggregator/index.html#trade-classifications).

```{r eval = FALSE, message=FALSE, warning=FALSE}
IM_DE_AT <- download_Comtrade(
  year = "2018",
  frequency = "annual",
  countries = "Germany",
  partners = "Austria",
  nomenclature = "HS2007",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "imports",
  type = "commodities",
  is.mirrorData = FALSE
)
IM_DE_AT %>%
  summarise(sum(trade_value_usd)) %>%
  pull() / mirrX_AT_DE %>%
    summarise(sum(trade_value_usd)) %>%
    pull()
#> [1] 1
```


## Generating mirrored World trad data


`comtradeRggregator` also offers simple options to download mirrored global trade data by downloading mirrored bilateral trade flows and aggregating it up to mirrored global trade data. In the simple example below, we download mirrored trade data for Austria. By setting `countries = "World"`, the package will create a mirrored data query for all global economies for the provided specification of year, trade flow, nomenclature, level of aggregation etc., and create a ready-to-use data frame.

In other words, to get the first mirrored export commodity flow for commodity group `01 - Live animals; animal products` in the query below, data from all countries that *import* this commodity from Austria are queried and aggregated. Then, when aggregating the final data set, the 'order' of the trade flows is reversed, i.e. the previously downloaded *imports* are labelled as *mirrored exports*. This reversal is highlighted in column `reporter`.

In elaborate cases such as the one above it is advisable to keep any temporarily downloaded data by setting `rm.temporaryFiles = FALSE`, and, ideally, to have access to an authenticated account; see [here](https://comtrade.un.org/data/doc/api/) for more information. 

```{r eval=FALSE, message=FALSE, warning=FALSE}
AT_mirr <- download_Comtrade(
  year = "2018",
  frequency = "annual",
  countries = "Austria",
  partners = "World",
  nomenclature = "HS2007",
  commodity = "ag2",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  build.Comtrade = TRUE,
)
AT_mirr
#> # A tibble: 363 x 8
#>    classification period trade_flow reporter         partner commodity_code commodity    ~ trade_value_usd
#>    <chr>           <int> <chr>      <chr>            <chr>   <chr>          <chr>        ~           <dbl>
#>  1 H3               2018 Export     Austria mirrored World   01             Live animals;~       146122111
#>  2 H3               2018 Export     Austria mirrored World   02             Meat and edib~      1279494900
#>  3 H3               2018 Export     Austria mirrored World   03             Fish and crus~        10067362
#>  4 H3               2018 Export     Austria mirrored World   04             Dairy produce~      1522685531
#>  5 H3               2018 Export     Austria mirrored World   05             Products of a~        79271602
#>  6 H3               2018 Export     Austria mirrored World   06             Live trees an~        26938801
#>  7 H3               2018 Export     Austria mirrored World   07             Edible vegeta~       142130800
#>  8 H3               2018 Export     Austria mirrored World   08             Edible fruit ~       209962703
#>  9 H3               2018 Export     Austria mirrored World   09             Coffee, tea, ~       116804431
#> 10 H3               2018 Export     Austria mirrored World   10             Cereals      ~       544971227
#> # ... with 353 more rows
```






## A more complex application

### The problem

Let's say you want to download monthly mirror trade data starting in 2019 at 
the HS 6-digit level for a handful of countries for which there is no trade data
available on [comtrade](https://comtrade.un.org/).

While there are multiple areas where both [comtrade](https://comtrade.un.org/) and 
[WITS](https://wits.worldbank.org/) offer similar functionality
to the `comtradeRggregator` package, this is a scenario where both alternatives
are tedious to work with:

- [WITS](https://wits.worldbank.org/) does offer a straight-forward integration of mirror data, but there are
 (at the time of writing this documentation in 2021) no options available to query 
 monthly data;
 
 - [Comtrade](https://comtrade.un.org/) does over the most recent and granular trade data, but there is no option to query mirror data.


Using `comtradeRggregator`, the this query can be put together with a couple
of line sof code.


### The implementation in `comtradeRggregator`



```{r, include = FALSE}
# Load token; do not show
my_token <- Sys.getenv("COMTRADE_TOKEN")
my_location <- "C:/Users/amann/Desktop/tmp"
```

First, we want to define list of countries for extraction. The two examples we 
use here are *Afghanistan* and the *Democratic Republic of the Congo*. 
To double-check the correct spelling of the country names we use 
`is.available_Comtrade()` with option `is.fuzzy = TRUE`. Below I run the check 
for the *Democratic Republic of the Congo* only. However, for large queries it
is advisable to double-check that all country names are correct before setting
up the actual data query. If you don't do that, and are unlucky in the sense that
a particular country name is spelled differently in 
[comtrade](https://comtrade.un.org/), your data query will return an error.

```{r message=FALSE, warning=FALSE}
is.available_Comtrade(
  is.contained = "Congo",
  is.fuzzy = TRUE,
  frequency = "annual",
  nomenclature = "HS",
  year = '2019'
)
```

The final list of countries we want to query is:

```{r}
country_list = c("Afghanistan", "Dem. Rep. of the Congo")
```


When setting up this more elaborate query, there are a few things we should
keep in mind:


- It makes sense to break up this query into two parts: a download for the years 2019 and 2020,
and a separate query for 2021. The reason for this is that at the time of writing 
this  tutorial, trade data for all 12 months was available for 2019 and 2020, 
but not for 2021 as this would make `download_Comtrade()`
throw an error message.


- It is advisable to keep any temporarily downloaded data by setting 
`rm.temporaryFiles = FALSE`. I also want to save the temporary files in a 
different location specified in variable `my_location`, so I also specify 
`location.temporaryFiles = my_location` in the query below. 

- In an ideal world, you will have access to an authenticated account 
[(see here)](https://comtrade.un.org/data/doc/api/). 
I do, so I'm also providing my [comtrade](https://comtrade.un.org/) 
authentication token as part of the query. 

- Note that for the first query we are not specifying 
the parameter `month` which, by default, will return data for all 12 months.
Also, note that monthly [comtrade](https://comtrade.un.org/) data is only available following the 
*HS combined nomenclature*.

- Finally, as we do not want to overwhelm the API, we reduce the number of 
extracted countries per sub-query to three, i.e. `ext_cnt  = 3`, instead of
the default of five, and increase the cool-down period between the separate 
queries from five to ten seconds between rounds by setting `speed = 10`. 
These last two steps are optional and only serve illustrative purposes.


With all of this out of the way, the final monthly mirror data query for 2019 and
2020 looks like this:

```{r eval=FALSE, message=FALSE, warning=FALSE}
mirr_part1 <- download_Comtrade(
  year = c("2019", "2020"),
  frequency = "monthly",
  countries = country_list,
  partners = "World",
  nomenclature = "HS",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  location.temporaryFiles = my_location,
  build.Comtrade = TRUE,
  token = my_token,
  ext_cnt  = 3,
  sleep = 10
)
```


For the 2021 query, we first want to find the last month of 2021 for which 
trade data is available. This makes sense because we do not want to run a query for
a month where no trade data is available as this would make `download_Comtrade()`
throw an error message.

To check the data availability, you can either use the 
`is.available_Comtrade()` function, or
the data file `Comtrade_DA` which I am doing below. As you can see, at the 
time of writing this tutorial, the most recent monthly trade data is July 2021.


```{r}
Comtrade_DA %>% 
  mutate(yr = substr(ps, 1, 4), mnth = substr(ps, 5,6)) %>% 
  filter(freq == "MONTHLY" & yr == '2021' ) %>%
  summarise(x = max(mnth))
```

The query for 2021 is identical to the first one with the exception of parameter
`month = c(1:7)`. 

```{r eval=FALSE, message=FALSE, warning=FALSE}
mirr_part2 <- download_Comtrade(
  year = c("2021"),
  frequency = "monthly",
  month = c(1:7),
  countries = country_list,
  partners = "World",
  nomenclature = "HS",
  commodity = "ag6",
  select.stats = "trade_value_usd",
  direction = "all",
  type = "commodities",
  is.mirrorData = TRUE,
  rm.temporaryFiles = FALSE,
  location.temporaryFiles = my_location,
  build.Comtrade = TRUE,
  token = my_token,
  ext_cnt  = 3,
  sleep = 10
)
```
